<span style="color:red;">Building a house with<br />Advanced Normalization Tools (*ANTs*)</span>

### 2015-05-09 evaluation/antsvfreesurfer2.Rmd
ENH:  More edits to the talk.
BUG:  Wrong filename.
ENH:  More stuff.

# ANTs brain morphology<br><br>[<small>"*Large-scale evaluation of ANTs and FreeSurfer cortical thickness measurements*"</small>](http://www.ncbi.nlm.nih.gov/pubmed/24879923)

## Cortical thickness studies:
```{r, results='asis', echo = FALSE}
conditions <- c( "Huntington’s disease", "schizophrenia", "bipolar disorder",
                 "Alzheimer’s disease", "frontotemporal dementia", "Parkinson’s disease",
                 "Williams syndrome", "multiple sclerosis", "autism", "migraines",
                 "chronic smoking", "alcoholism", "cocaine addiction", "Tourette syndrome in children",
                 "scoliosis in female adolescents", "early-onset blindness", "chronic pancreatitis",
                 "obsessive-compulsive disorder", "ADHD", "obesity", "heritable depression",
                 "elderly depression", "age", "gender", "untreated male-to-female transsexuality",
                 "handedness", "intelligence", "athletic ability", "meditative practices", "musical ability",
                 "tendency toward criminality", "childhood sexual abuse in female adolescents",
                 "Tetris-playing ability in female adolescents", "traumatic brain injury" )

@@

## *ANTs*-related background

* Sandy, Brian, and <span style="color:red;">DiReCT</span>
  + used other tools such as FMRIB's Automated Segmentation Tool (FAST)
  + `KellySlater.cxx`
  * scripts "adequate for Penn data"
* `KellyKapowski.cxx` and `antsCorticalThickness.sh`
* _"Let's evaluate on open data."_
* and, eventually, _"Let's compare with FreeSurfer."_

## The *ANTs* structural brain mapping workflow

@@

## Basic components of the pipeline

@@

## Template building

@@

## Template building (cont.)

@@

## Brain extraction comparison<br>*ANTs* vs. FreeSurfer

@@

## Brain segmentation

@@

Randomly selected healthy individuals.  Atropos gets good performance across ages.

## Cortical thickness estimation

@@

## _But without ground truth, how does one evaluate the pipeline?_

* Use age and gender (demographic information) in training/prediction paradigm
* Public data sets (IXI, NKI, Oasis, MMRR)
* $>$ 1200 subjects, age 7 to over 90 years old

## Prediction from cortical thickness data:<br>gender and age

@@

## Regional importance comparison

@@

## Regional importance comparison (cont.)

<div align="center"><img height="550" src="./evaluation/figures/antsvfreesurfer_regionalPlots.png" frameborder="0"></img></div>

# Cortical thickness extensions

>>>>

# ANTs brain morphology<br>[<small>"*Large-scale evaluation of ANTs and FreeSurfer cortical thickness measurements*"</small>](http://www.ncbi.nlm.nih.gov/pubmed/24879923)

##

__Cortical thickness studies__

```{r, results='asis', echo = FALSE}
conditions <- c( "Huntington’s disease", "schizophrenia", "bipolar disorder",
                 "Alzheimer’s disease", "frontotemporal dementia", "Parkinson’s disease",
                 "Williams syndrome", "multiple sclerosis", "autism", "migraines",
                 "chronic smoking", "alcoholism", "cocaine addiction", "Tourette syndrome in children",
                 "scoliosis in female adolescents", "early-onset blindness", "chronic pancreatitis",
                 "obsessive-compulsive disorder", "ADHD", "obesity", "heritable depression",
                 "elderly depression", "age", "gender", "untreated male-to-female transsexuality",
                 "handedness", "intelligence", "athletic ability", "meditative practices", "musical ability",
                 "tendency toward criminality", "childhood sexual abuse in female adolescents",
                 "Tetris-playing ability in female adolescents", "traumatic brain injury" )

@@

##

__ANTs-related background__

* Sandy, Brian, and <span style="color:red;">DiReCT</span>
    + used other tools such as FMRIB Automated Segmentation Tool (FAST)
    + `KellySlater.cxx`
* pipedream scripts "adequate for Penn data."
* `KellyKapowski.cxx` and `antsCorticalThickness.sh`
* _"Let's evaluate on open data."_
* and, eventually, _"Let's compare with FreeSurfer."_

##

__The ANTs structural brain mapping workflow__

@@

##

__Basic components of the pipeline__

@@

##

__Template building__

@@

##

__Template building__ (cont.)

@@

##

__Brain extraction comparison ANTs vs. FreeSurfer__

@@

##

__Brain segmentation__

@@

Randomly selected healthy individuals.  ```Atropos``` gets good performance across ages.

##

__Cortical thickness estimation__

@@

## <span style="color:red;">_But without ground truth, how does one evaluate the pipeline?_</span>

##

<div align="center"><img height="400" src="./evaluation/figures/ageDistribution.png" frameborder="0"></img></div>

* Use age and gender (demographic information) in training/prediction paradigm
* Public data sets (IXI, NKI, OASIS, MMRR)
* $>$ 1200 subjects, age 4 to over 94 years old

##

__Prediction from cortical thickness data:<br>gender and age__

@@

##

__Age prediction per site__

<div align="center"><img height="350" src="./evaluation/figures/agePredictionPerSite.png" frameborder="0"></img></div>

##

__Regional importance comparison__

@@

##

__Regional measurements__
<div align="center"><img height="550" src="./evaluation/figures/antsvfreesurfer_regionalPlots.png" frameborder="0"></img></div>

### 2015-05-09 installation/antsr_installation.Rmd
ENH:  More stuff.

> antsrDependencies <- c( "Rcpp", "tools", "methods", "magrittr" )
> install.packages( pkgs = antsrDependencies, dependencies = TRUE )
~~~~~~~~~~~~~~~~~~

>>>>

> antsrDependencies <- c( "Rcpp", "tools", "methods", "magrittr" )
> install.packages( pkgs = antsrDependencies, dependencies = TRUE )
> moreAntsrDependencies <- c( "igraph", "randomForest", "misc3d", "rgl", "mFilter", "fastICA", "fpc", "knitr", "rmarkdown", "pixmap", "png", "signal", "visreg" )
> install.packages( moreAntsrDependencies )

~~~~~~~~~~~~~~~~~~

### 2015-05-09 lesions/lesions.Rmd
ENH:  More stuff.

## antsCorticalThickness.sh is adaptable!

1. Register subject (or single subject template) to normal template.

@@

3. Create additional "lesion" prior, i.e. ```SmoothImage 3 ${lesionMask} 1 ${lesionPrior} 1```.

@@

## Modified template spatial priors

![](lesions/figures/lesionPrior.png)

## antsCorticalThickness.sh using lesion prior

![](lesions/figures/lesionSegmentation.png)

Only change to the command call is an additional '-c WM[7]' which means "combine
the 7<sup>th</sup> prior, i.e. lesion, to the white matter for cortical thickness calculation."

>>>>

##
```antsCorticalThickness.sh``` is adaptable!

1. Register subject (or single subject template) to normal template with priors.

@@

3. Create additional "lesion" prior, i.e.

~~~~~~~~~~~~~~~~~~~~~~~
SmoothImage 3 ${lesionMask} 1 ${lesionPrior} 1
~~~~~~~~~~~~~~~~~~~~~~~

@@

##

__Modified template spatial priors__

<div align="center"><img height="450" src="lesions/figures/lesionPrior.png" frameborder="0"></img></div>

##

```antsCorticalThickness.sh``` using lesion prior

<div align="center"><img height="450" src="lesions/figures/lesionSegmentation.png" frameborder="0"></img></div>

<!--
Only change to the command call is an additional '-c WM[7]' which means "combine
the 7<sup>th</sup> prior, i.e. lesion, to the white matter for cortical thickness calculation."
-->
